---
title: 'How did economic developments influence inaugural speeches? '
output:
  html_notebook: default
  html_document: default
  pdf_document: default
---
### Introduction

Economic conditions are vital for a country's development,that' why every president talks at length about them during their inaugural speeches.However,there may exist a couple of differences in aspects of their speeches between prosperous years and depression years.Unfortunately,yet it's only our guessing.In this analysis,we'll see how economic development makes effects on presidents' inaugual speeches.Besides,we compare two presidents, Roosevelt and Obama, both of whom experienced depression,and see what convergence and divergence are in their speeches.

## Step 0: check and install needed packages. Load the libraries and functions
```{r, message=FALSE, warning=FALSE}
packages.used=c("tm", "wordcloud", "RColorBrewer", 
                "dplyr", "tidytext")
# check packages that need to be installed.
packages.needed=setdiff(packages.used, 
                        intersect(installed.packages()[,1], 
                                  packages.used))
# install additional packages
if(length(packages.needed)>0){
  install.packages(packages.needed, dependencies = TRUE,
                   repos='http://cran.us.r-project.org')
}


library("rvest")
library("tibble")
library("qdap")
library("sentimentr")
library("gplots")
library("dplyr")
library("tm")
library("syuzhet")
library("factoextra")
library("scales")
library("RColorBrewer")
library("RANN")
library("topicmodels")
library("tidytext")
library("wordcloud")
library("beeswarm")


source("/Users/1/Desktop/ADS/Project 1/lib/speechFuncs.R")
source("/Users/1/Desktop/ADS/Project 1/lib/plotstacked.R")
```
## Step 1: scrap speech URLs from website and import GDP data from local file 
```{r}
### Inauguaral speeches
main.page <- read_html(x = "http://www.presidency.ucsb.edu/inaugurals.php")
# Get link URLs
inaug=f.speechlinks(main.page)
inaug=inaug[-nrow(inaug),] # remove the last line, irrelevant due to error.
inaug.list=read.csv("/Users/1/Desktop/ADS/Project 1/data/inauglist.csv", stringsAsFactors = FALSE)
inaug.list=cbind(inaug.list,inaug)
inaug.list$type="inaug"


# Import economic data
gdp=read.csv("/Users/1/Desktop/ADS/Project 1/data/GDP.csv",skip=1,header=TRUE)
gdp=gdp[,-2]
names(gdp) <- c("year","growth")
gdp$year=as.Date(paste(gdp$year,"-12-31",sep=""))


inaug.list$Date=as.Date(inaug[,1], format="%B %e, %Y")

inaug.list=filter(inaug.list,Date>gdp$year[1])
inaug.list$growth=NA
for (i in seq(nrow(inaug.list)-1)){
  years<-filter(gdp,year>inaug.list$Date[i] & year<=inaug.list$Date[i+1])
  inaug.list$growth[i]=mean(years$growth)
}


# Loop over each row in inaug.list
inaug.list$fulltext=NA
for(i in seq(nrow(inaug.list))) {
  text <- read_html(inaug.list$urls[i]) %>% # load the page
    html_nodes(".displaytext") %>% # isloate the text
    html_text() # get the text
  inaug.list$fulltext[i]=text
  # Create the file name
  filename <- paste0("/Users/1/Desktop/ADS/Project 1/data/InauguralSpeeches/", 
                     inaug.list$type[i], 
                     inaug.list$File[i], "-", 
                     inaug.list$Term[i], ".txt")
  sink(file = filename) %>% # open file to write 
  cat(text)  # write the file
  sink() # close the file
}
```

In this step,only speeches given after 1929 are included because of the availability of economic data.

## Step 2: data Processing
```{r}
folder.path="/Users/1/Desktop/ADS/Project 1/data/InauguralSpeeches/"
speeches=list.files(path = folder.path, pattern = "*.txt")
prex.out=substr(speeches, 6, nchar(speeches)-4)

ff.all<-Corpus(DirSource(folder.path))

ff.all<-tm_map(ff.all, stripWhitespace)
ff.all<-tm_map(ff.all, content_transformer(tolower))
ff.all<-tm_map(ff.all, removeWords, stopwords("english"))
ff.all<-tm_map(ff.all, removeWords, character(0))
ff.all<-tm_map(ff.all, removePunctuation)
ff.all<-tm_map(ff.all, stemDocument)
```

For the speeches, we removed extra white space, converted all letters to the lower case, removed stop words, removed empty words due to formatting errors, and removed punctuation.

## Step 3: data analysis --- topic modeling

Gengerated document-term matrices.
```{r}
dtm <- DocumentTermMatrix(ff.all)
#convert rownames to filenames
rownames(dtm) <- paste0(inaug.list$type, inaug.list$File,"_",inaug.list$Term)
rowTotals <- apply(dtm , 1, sum) #Find the sum of words in each Document 
dtm  <- dtm[rowTotals> 0, ]
```

Run LDA
```{r}
#Set parameters for Gibbs sampling
burnin <- 4000
iter <- 1000
thin <- 500
seed <-list(2003,2,68,100001,765)
nstart <- 5
best <- TRUE

# Number of topics
k <- 15

# Run LDA using Gibbs sampling
ldaOut <-LDA(dtm, k, method="Gibbs", control=list(nstart=nstart, 
                                                 seed = seed, best=best,
                                                 burnin = burnin, iter=iter,
                                                 thin=thin))
```

For each topic,the 15 most common terms were extracted,and the 10 most salient terms as well.However,only the first 5 are displayed.
```{r}
# write out results
# docs to topics
ldaOut.topics <- as.matrix(topics(ldaOut))
write.csv(ldaOut.topics,file=paste("/Users/1/Desktop/ADS/Project 1/output/LDAGibbs",k,"DocsToTopics.csv"))

#top 6 terms in each topic
ldaOut.terms <- as.matrix(terms(ldaOut,15))
write.csv(ldaOut.terms,file=paste("/Users/1/Desktop/ADS/Project 1/output/LDAGibbs",k,"TopicsToTerms.csv"))

#probabilities associated with each topic assignment
topicProbabilities <- as.data.frame(ldaOut@gamma)
write.csv(topicProbabilities,file=paste("/Users/1/Desktop/ADS/Project 1/output/LDAGibbs",k,"TopicProbabilities.csv"))
# each row represents a document, and rowSum equals 1

# logarithmized parameters of the word distribution for each topic.
terms.beta=ldaOut@beta
terms.beta=scale(terms.beta)
topics.terms=NULL
for(i in 1:k){
  topics.terms=cbind(topics.terms, ldaOut@terms[order(terms.beta[i,], decreasing = TRUE)[1:10]]) # select 10 keywords for each topic
}

write.csv(topics.terms,file=paste("/Users/1/Desktop/ADS/Project 1/output/LDAGibbs",k,"TopicsKeywords.csv"))
```

Summarized each topic with one word topics.hash=c("Tax", "Defense", "America", "Economy", "America", "Prosperity", "Leadership", "Reform", "Liberty", "International", "Nation", "Freedom", "Democracy", "Misc", "Power")
```{r}
topics.hash=c("Tax", "Defense", "Crisis", "Economy", "America", "Prosperity", "Leadership", "Reform", "Liberty", "International", "Nation", "Freedom", "Democracy", "Misc", "Power")
inaug.list$ldatopic=as.vector(ldaOut.topics)
inaug.list$ldahash=topics.hash[ldaOut.topics]

colnames(topicProbabilities)=topics.hash
inaug.list.df=cbind(inaug.list, topicProbabilities)
```

Economic condition changes over time
```{r}
par(mar=c(1,1,2,0))
plot(inaug.list.df$Date,inaug.list.df$growth,ylab="GDP Growth %",xlab="Year",main="Economic condition Over Time",type="l",col=2)
```

Different topics were chosen during different periods of time
```{r, message=FALSE, warning=FALSE}
par(mar=c(1,1,1,1)) 

inaug.list.df$growth[nrow(inaug.list.df)]=0
inaug.list.df$growth2=NULL
inaug.list.df$growth2<-cut(inaug.list.df$growth,breaks=5)

topic.summary2=tbl_df(inaug.list.df)%>%
              group_by(growth2)%>%
              select(growth2, Tax:Power)%>%
              summarise_each(funs(mean))
topic.summary2=as.data.frame(topic.summary2)
rownames(topic.summary2)=topic.summary2[,1]

topic.plot=c(2:6,7,8,13)

heatmap.2(as.matrix(topic.summary2[,topic.plot+1]), 
          scale = "column", key=F, 
          col = bluered(100),
          cexRow = 0.9, cexCol = 0.9, margins = c(8, 8),
          trace = "none", density.info = "none")
```

It's obvious that presidents talked more about Reform, Prosperity,Derense during more prosperous years,while America and Economy were mentioned more frequently when the economic enviroment was depressed.
```{r}
inaug.list.df[which.max(inaug.list.df$growth),]$President
inaug.list.df[which.min(inaug.list.df$growth),]$President
```

Roosevelt is the president in the year with both highest and lowest GDP growth rate. 

## Step 3: data analysis --- wordcloud comparison

There were two devestating economic depressions in American history.One is the Great Depression in 1929,when Roosevelt was president of U.S.Another one happened in 2008,Obama became president at that point.So many news reports talked about the similarities between them.Let's make it more detailed with wordcloud and emotion analysis. 

This is Roosevelt's wordcloud.
```{r}
tdm.all<-TermDocumentMatrix(ff.all)
tdm.tidy=tidy(tdm.all)
tdm.all = subset(tdm.tidy,document == 'inaugFranklinDRoosevelt-1.txt'|document =='inaugFranklinDRoosevelt-2.txt')
#tdm.overall=summarise(group_by(tdm.tidy, term), sum(count))
wordcloud(tdm.all$term, tdm.all$count,
          scale=c(2,0.5),
          max.words=100,
          min.freq=1,
          random.order=FALSE,
          rot.per=0.3,
          use.r.layout=T,
          random.color=FALSE,
          colors=brewer.pal(9,"Blues"))
```

This is Obama's wordcloud.
```{r}
tdm.all<-TermDocumentMatrix(ff.all)
tdm.tidy=tidy(tdm.all)
tdm.all = subset(tdm.tidy,document == 'inaugBarackObama-1.txt'|document =='inaugBarackObama-2.txt')
#tdm.overall=summarise(group_by(tdm.tidy, term), sum(count))
wordcloud(tdm.all$term, tdm.all$count,
          scale=c(2,0.5),
          max.words=100,
          min.freq=1,
          random.order=FALSE,
          rot.per=0.3,
          use.r.layout=T,
          random.color=FALSE,
          colors=brewer.pal(9,"Blues"))
```

From the two pictures we got,we can easily find both Roosevelt's and Obama's focus on national development and human rights.Besides,although they used economic words like work, money etc,powerful words such as effort,million,duty were used more often by Roosevelt than Obama,as shown in picture one.The Great Depression was worse than the financial crisis that happened in 2008,that's why Roosevelt used more emotional and economic words compared to Obama.It agrees with our original thoughts. 

## Step 3: data analysis --- emotion analysis
```{r, message=FALSE, warning=FALSE}
sentence.list=NULL
for(i in 1:nrow(inaug.list)){
  sentences=sent_detect(inaug.list$fulltext[i],
                        endmarks = c("?", ".", "!", "|",";"))
  if(length(sentences)>0){
    emotions=get_nrc_sentiment(sentences)
    word.count=word_count(sentences)
    # colnames(emotions)=paste0("emo.", colnames(emotions))
    # in case the word counts are zeros?
    emotions=diag(1/(word.count+0.01))%*%as.matrix(emotions)
    sentence.list=rbind(sentence.list, 
                        cbind(inaug.list[i,-ncol(inaug.list)],
                              sentences=as.character(sentences), 
                              word.count,
                              emotions,
                              sent.id=1:length(sentences)
                              )
    )
  }
}
sentence.list=
  sentence.list%>%
  filter(!is.na(word.count))


sel.comparison2=c("FranklinDRoosevelt")
sentence.list1=filter(sentence.list, 
                       File%in%sel.comparison2)
heatmap.2(cor(sentence.list1%>%select(anger:trust)), 
          scale = "none", 
          col = bluered(100), , margin=c(6, 6), key=F,
          trace = "none", density.info = "none",main="Roosevelt's heatmap")
```

```{r}
sel.comparison3=c("BarackObama")
sentence.list2=filter(sentence.list, 
                       File%in%sel.comparison3)
heatmap.2(cor(sentence.list2%>%select(anger:trust)), 
          scale = "none", 
          col = bluered(100), , margin=c(6, 6), key=F,
          trace = "none", density.info = "none",main="Obama's heatmap")
```

Apparently,they felt sad and disgusted about America's situation,but at the same time,they also had positive emotion like trust in the future.What makes Roosevelt and Obama different is Roosevelt had a higher proportion of negative emotions,while Obama kept almost perfect balance between positive emotions and negatie emotions.It perfectly matches what we found in the wordcloud part.Roosevelt faced more challenges about enhancing American economy.

### Conclusion

So far,we simply analyzed how economic development influenced presidents' inaugual speeches,and looked at the similarities and differences between Roosevelt and Obama by text modeling,wordcloud and emotion analysis.We can conclude that:

First,during prosperous years,these presidents talked more about Reform, Prosperity and Derense,while America and Economy were mentioned more frequently when the economic enviroment was depressed.We can see that these presidents attached more importance on economic development when GDP growth rate was low.

Moreover,Roosevelt became the president of the U.S. while the Great Depression was happening.However,with a lot of hard work he made the GDP increase tremendously,even the highest GDP in the next few years.This is a strong proof for how efficient and powerful Roosevelt's policy was.

Last but not least,we noticed Obama shared some similiries with Roosevelt.Both of them became president when the economic condition in US was depressed.Both Roosevelt and Obama focused on national development and human rights.And they apparently felt sad and disgusted about American situation.But Roosevelt had a higher proportion of negative emotions,while Obama kept almost perfect balance between positive emotions and negatie emotions.Also,Roosevelt used more economic words than Obama.Powerful words such as effort,million,duty were used by Roosevelt.It's known that the Great Depression was worse than the financial crisis that happened in 2008,maybe that's why Roosevelt used more emotional and economic words compared to Obama.